The genomic impacts of drift and selection for hybrid performance in maize

### 2015-08-19 response.tex

I was quite intrigued by the difference in distance between cycles 4 and then 8-16.  The authors suggest that this could be due to some differences in genetic maintenance of the populations.  I am curious if there were any particular environmental influences during that time period that may have led to stronger selection or if there were any differences in agronomic practices that have occurred during the long-term experiment (ie irrigation, fertilization, plant density, etc).  it would be helpful to know if any of these factors may have contributed.

\response{The major difference is the change to 20 (instead of 10) lines at cycle 8. Machine harvest was introduced at cycle 6, but we have no reason to believe this would have dramatically changed diversity patterns. From our reading of the literature and consultation with breeders who have run the experiment we are not aware of any other significant changes. }

It was a bit surprising to me that there was no differentiation among the founder inbreds of the two populations.  If PCA was run only on the founders (without any other lines) do you get separation of the two groups or not?

\response{Projections of all lines onto a PCA of the founders shows no differentiation corresponding to BSSS and BSCB1. The main eigenvector only explains $\approx 6\%$ of the variation in the founders and appears to separate Reid and Lancaster lines.  The BSSS and BSCB1 plants form two clusters near zero along this axis. This figure has been included in the supplement.  }

@@

Also related to the simulations, there are multiple references in the text to empirical results as "deviation from 
simulation" (e.g., p 15) when I suspect what the authors wish to say is that the results differ from neutral 
expectations based on simulations.

\response{We have removed the phrase ``deviation from simulation'' throughout, replacing it with phrases mentioning differences from neutral simulations.}

The other major weakness of the paper is that relatively little background on the experimental populations 
is provided. There is a long list of citations of prior work on this experimental population, but little 
exposition.

\response{We have added a paragraph in the introduction explaining the RRS method and populations in increased detail.\jri{this enough or do we add more elsewhere?}}

@@

A better presentation of the specific target of selection is needed (I infer that selection is on 
improved yield through increased hybrid performance). Ultimately in the Discussion, there is additional 
discussion of the phenotypes, but the presentation is minimal (again pointing to multiple citations) and 
provides little insight into the nature of the selection observed. Better discussion of phenotypes potentially 
targeted by selection is essential if the manuscript is to end, as it does, with a call for further dissection of 
targets of selection in this population.

\response{We have noted in the new introductory paragraph that yield was the main focus of selection.\jri{add other phenotypes elsewhere?}}

### 2015-08-19 main.tex
jode text

Multiple studies with molecular markers have indicated that these heterotic groups have diverged genetically over time to become highly structured and isolated populations, resulting in a dramatic restructuring of population genetic variation \citep{duvick2004long, ho2005extent, feng2006temporal}. 
Advances in high throughput genotyping and the development of a maize reference genome now enable the observation of maize population structure at high marker density across the whole genome \citep{ganal2011a-large,chia2012maize}. 
These studies have examined a broad spectrum of germplasm at various points in the history of maize to search for the signals of population structure and artificial selection \citep{Hufford2012b, van2012historical,jiao2012genome}. 
Although selective sweeps remaining from domestication are clearly visible, the impact of selection during modern breeding appears comparatively small in terms of its impact on genomic diversity \citep{Hufford2012b,van2012historical}, despite steady, heritable improvement in phenotype \citep{duvick2005contribution}. 
The lack of distinct selection signals from modern breeding may be due to population-specific selection; this interpretation is supported by the measured success in identifying targets of selection in individual experimental populations selected for a limited number of phenotypes \citep{hirsch2014insights, beissinger2014genome}.

Here we take a detailed look at the dynamics of genome-wide patterns of diversity over time within an individual breeding program. 
The Iowa Stiff Stalk Synthetic (BSSS) and the Iowa Corn Borer Synthetic No. 1 (BSCB1) Reciprocal Recurrent Selection Program of the USDA-ARS at Ames, Iowa (hereafter referred to as the Iowa RRS) represents one of the best-documented public experiments on selection for combining ability and hybrid performance. 
BSSS and BSCB1 have been recurrently selected for improved cross-population hybrids \citep{penny1971twenty}. 
\jri{Jode/Justin add text explaining recurrent selection and populations}
This model of selection, named reciprocal recurrent selection, provides the generalized model for strategies commonly used in commercial maize breeding \citep{comstock1949breeding, duvick2004long}. 
The Iowa RRS experiment proves especially relevant because lines derived from the BSSS population have had a major impact upon the development of commercial hybrids \citep{duvick2004long, darrah1986}, the formation of modern heterotic groups \citep{troyer1999background, senior1998utility}, and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.
\jri{something about heterosis}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRO

>>>>

\rev{While the founders of these heterotic groups were not initially differentiated, m}ultiple studies with molecular markers have indicated that these heterotic groups have diverged genetically over time to become highly structured and isolated populations, resulting in a dramatic restructuring of population genetic variation \citep{duvick2004long, ho2005extent, feng2006temporal}. 
Advances in high throughput genotyping and the development of a maize reference genome now enable the observation of maize population structure at high marker density across the whole genome \citep{ganal2011a-large,chia2012maize}. 
These studies have examined a broad spectrum of germplasm at various points in the history of maize to search for the signals of population structure and artificial selection \citep{Hufford2012b, van2012historical,jiao2012genome}. 
Although selective sweeps remaining from domestication are clearly visible, the impact of selection during modern breeding appears comparatively small in terms of its impact on genomic diversity \citep{Hufford2012b,van2012historical}, despite steady, heritable improvement in phenotype \citep{duvick2005contribution}. 
The lack of distinct selection signals \rev{in broad germplasm collections may be due to population-specific selection within germplasm subgroups; in this case selection should be easier to detect in populations maintained in an individual program.} This \rev{possibliity} is supported by the measured success in identifying targets of selection in individual experimental populations \rev{under directional selection for specific phenotypes such as seed size and ear number} \citep{hirsch2014insights, beissinger2014genome}.

\rev{In this study, we examine a different closed experiment whose phenotypic target and selection method --- reciprocal recurrent aelection --- closely mirrors methods used in the generation of modern maize hybrids \citep{comstock1949breeding, duvick2004long}. 
Proposed by \citet{comstock1949breeding}, reciprocal recurrent selection is a procedure to simultaneously improve two populations for their ability to produce high yielding hybrids when individuals from the two populations are crossed. 
In this method, the two populations are kept separate and lines from each population are selected based on the phenotype of their cross-population hybrids. 
Selected lines within each population are then intermated to generate lines for the next cycle. 
In this way, the two populations simultaneously evolve improved combining ability with one another. 
The USDA-ARS at Ames, Iowa has conducted a reciprocal recurrent selection program (hereafter referred to as the Iowa RRS) with the Iowa Stiff Stalk Synthetic (BSSS) and Iowa Corn Borer Synthetic No. 1 (BSCB1) populations for 18 cycles of selection\citep{penny1971twenty}. 
This program represents one of the best-documented public experiments on selection for combining ability and hybrid performance. 
Extensive records on population sizes, breeding methods, selection differentials, and seed are available for all cycles of selection. 
This set of resources on a closed population selected for hybrid yield make the Iowa RRS an ideal test case for the study of the genomic impact of intensive hybrid breeding in maize.} 
The Iowa RRS experiment \rev{provides additional historical relevance} because lines derived from the BSSS population have had a major impact on the development of commercial hybrids \citep{duvick2004long, darrah1986}, the formation of modern heterotic groups \citep{troyer1999background, senior1998utility}, and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.
\jri{something about heterosis} \jpg{what are you thinking for heterosis here?}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRO
